---
title: "Binomial Representation Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5)
```

## Setup

```{r packages}
library(tidyverse)
library(brms)
library(tidybayes)
library(patchwork)
library(broom)
```

```{r load-data}
MODEL_LEVELS <- c("125m", "350m", "1.3b", "olmo-1b", "olmo-7b")

# Frequency data: BabyLM corpus for small models, Dolma (infini-gram) for OLMo
babylm_freqs <- read_csv("../Data/babylm_corpus_freqs.csv") %>%
  select(phrase_AB, overall_freq, rel_freq)

infinigram_freqs <- read_csv("../Data/infinigram_freqs.csv") %>%
  select(phrase_AB, overall_freq, rel_freq)

# Outcome data
proc_raw <- read_csv("../results/binomial_representations.csv") %>%
  select(-any_of(c("overall_freq", "rel_freq")))

cos_raw  <- read_csv("../Data/cosine_similarity.csv")
comp_raw <- read_csv("../Data/compositional_similarity.csv")
```

```{r helpers}
# Join the right frequency table based on model_size
join_freqs <- function(df) {
  bind_rows(
    df %>%
      filter(model_size %in% c("125m", "350m", "1.3b")) %>%
      left_join(babylm_freqs, by = "phrase_AB"),
    df %>%
      filter(model_size %in% c("olmo-1b", "olmo-7b")) %>%
      left_join(infinigram_freqs, by = "phrase_AB")
  )
}

# Keep only the final checkpoint × last layer per model
filter_final <- function(df) {
  df %>%
    group_by(model_size) %>%
    filter(tokens == max(tokens), layer == max(layer)) %>%
    ungroup()
}

sc <- function(x) as.vector(scale(x))

# Shared brms priors for all Gaussian models
base_priors <- c(
  prior(normal(0, 1),   class = b),
  prior(normal(0, 1),   class = Intercept),
  prior(exponential(1), class = sigma)
)

# Fit linear vs quadratic per group; select by AIC (delta > 2 → quadratic).
# Returns a prediction df: <x_col>, .fit, .lwr, .upr, + group columns.
fit_smart_smooth <- function(data, x_col, y_col,
                              group_cols = "model_size",
                              delta_aic  = 2) {
  data %>%
    filter(!is.na(.data[[x_col]]), !is.na(.data[[y_col]])) %>%
    group_by(across(all_of(group_cols))) %>%
    group_modify(function(d, key) {
      x   <- d[[x_col]]; y <- d[[y_col]]
      if (length(x) < 6) return(tibble())
      m1  <- lm(y ~ x)
      m2  <- lm(y ~ poly(x, 2))
      mod <- if (AIC(m2) < AIC(m1) - delta_aic) m2 else m1
      xs  <- seq(min(x), max(x), length.out = 200)
      p   <- predict(mod, newdata = data.frame(x = xs), interval = "confidence")
      out <- tibble(.fit = p[, "fit"], .lwr = p[, "lwr"], .upr = p[, "upr"])
      out[[x_col]] <- xs
      out
    }) %>%
    ungroup()
}
```

---

## Procrustes Distance

Procrustes distance between the AB and BA representation clouds at the final
checkpoint, last layer. Lower distance = model treats both orderings similarly;
higher distance = ordering changes the representation substantially.

**Predictors:**
- `log_freq`: log overall corpus frequency
- `ord_dom = |RelFreq − 0.5|`: ordering dominance (0 = equally frequent in both orders, 0.5 = only one order attested)

```{r proc-prep}
proc <- proc_raw %>%
  mutate(model_size = factor(model_size, levels = MODEL_LEVELS)) %>%
  filter_final() %>%
  join_freqs() %>%
  filter(!is.na(overall_freq), overall_freq > 0,
         !is.na(rel_freq), !is.na(procrustes_dist)) %>%
  mutate(
    log_freq     = log(overall_freq),
    log_freq_sc  = sc(log_freq),
    ord_dom      = abs(rel_freq - 0.5),
    ord_dom_sc   = sc(ord_dom),
    proc_sc      = sc(procrustes_dist),
    log_freq_bin = cut(log_freq, breaks = 3,
                       labels = c("Low freq", "Mid freq", "High freq")),
    ord_dom_bin  = factor(
      case_when(ntile(ord_dom, 3) == 1 ~ "Balanced",
                ntile(ord_dom, 3) == 2 ~ "Moderate",
                ntile(ord_dom, 3) == 3 ~ "Dominant"),
      levels = c("Balanced", "Moderate", "Dominant")
    ),
    model_size   = factor(model_size, levels = MODEL_LEVELS)
  )

proc_sizes <- levels(droplevels(proc$model_size))
cat("Models:", paste(proc_sizes, collapse = ", "), "\n")
cat("N per model:\n"); print(count(proc, model_size))
```

### Visualizations

```{r proc-viz-freq, fig.height=5}
.sm <- fit_smart_smooth(proc, "log_freq", "procrustes_dist")
proc %>%
  ggplot(aes(x = log_freq, y = procrustes_dist, colour = ord_dom)) +
  geom_point(alpha = 0.4, size = 0.9) +
  geom_ribbon(data = .sm, aes(x = log_freq, ymin = .lwr, ymax = .upr),
              inherit.aes = FALSE, alpha = 0.15, fill = "grey40") +
  geom_line(data = .sm, aes(x = log_freq, y = .fit),
            inherit.aes = FALSE, colour = "black", linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1, scales = "free_x") +
  scale_colour_viridis_c(name = "Ordering dominance\n|RelFreq − 0.5|") +
  labs(x = "log(overall frequency)", y = "Procrustes distance",
       title = "Procrustes distance vs frequency (final checkpoint, last layer)") +
  theme_minimal() +
  theme(legend.position = "right")
```

```{r proc-viz-ord, fig.height=5}
.sm <- fit_smart_smooth(proc, "ord_dom", "procrustes_dist",
                        group_cols = c("model_size", "log_freq_bin"))
proc %>%
  ggplot(aes(x = ord_dom, y = procrustes_dist, colour = log_freq_bin)) +
  geom_point(alpha = 0.4, size = 0.9) +
  geom_ribbon(data = .sm, aes(x = ord_dom, ymin = .lwr, ymax = .upr,
                               fill = log_freq_bin),
              inherit.aes = FALSE, alpha = 0.2) +
  geom_line(data = .sm, aes(x = ord_dom, y = .fit, colour = log_freq_bin),
            inherit.aes = FALSE, linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1) +
  scale_colour_brewer(palette = "Blues", aesthetics = c("colour", "fill"),
                      name = "Frequency tertile") +
  labs(x = "Ordering dominance |RelFreq − 0.5|", y = "Procrustes distance",
       title = "Procrustes distance vs ordering dominance") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r proc-viz-interaction, fig.height=5}
.d  <- proc %>% filter(!is.na(ord_dom_bin))
.sm <- fit_smart_smooth(.d, "log_freq", "procrustes_dist",
                        group_cols = c("model_size", "ord_dom_bin"))
.d %>%
  ggplot(aes(x = log_freq, y = procrustes_dist,
             colour = ord_dom_bin, fill = ord_dom_bin)) +
  geom_point(alpha = 0.2, size = 0.7) +
  geom_ribbon(data = .sm, aes(x = log_freq, ymin = .lwr, ymax = .upr,
                               fill = ord_dom_bin),
              inherit.aes = FALSE, alpha = 0.2) +
  geom_line(data = .sm, aes(x = log_freq, y = .fit, colour = ord_dom_bin),
            inherit.aes = FALSE, linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1, scales = "free_x") +
  scale_colour_brewer(palette = "Set2", aesthetics = c("colour", "fill"),
                      name = "Ordering dominance") +
  labs(x = "log(overall frequency)", y = "Procrustes distance",
       title = "Frequency × ordering dominance interaction (Procrustes)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### OLS Coefficients

```{r proc-lm-coef, fig.height=4}
proc_sizes %>%
  set_names() %>%
  map_dfr(function(size) {
    lm(procrustes_dist ~ log_freq_sc * ord_dom_sc,
       data = proc %>% filter(model_size == size)) %>%
      tidy(conf.int = TRUE) %>%
      filter(term != "(Intercept)")
  }, .id = "model_size") %>%
  mutate(
    model_size = factor(model_size, levels = MODEL_LEVELS),
    term = recode(term,
                  log_freq_sc              = "log freq",
                  ord_dom_sc               = "ordering dominance",
                  `log_freq_sc:ord_dom_sc` = "freq × dom")
  ) %>%
  ggplot(aes(x = estimate, xmin = conf.low, xmax = conf.high,
             y = term, colour = model_size)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_pointrange(position = position_dodge(width = 0.5), linewidth = 0.8) +
  scale_colour_brewer(palette = "Set1", name = "Model") +
  labs(x = "OLS estimate (95% CI)", y = NULL,
       title = "Procrustes distance: OLS coefficients per model") +
  theme_minimal()
```

### brms Models

```{r proc-brms-fit, results="hide"}
proc_brms <- proc_sizes %>%
  set_names() %>%
  map(function(size) {
    d <- proc %>%
      filter(model_size == size) %>%
      mutate(
        proc_sc     = sc(procrustes_dist),
        log_freq_sc = sc(log_freq),
        ord_dom_sc  = sc(ord_dom)
      )
    brm(
      proc_sc ~ log_freq_sc * ord_dom_sc,
      data   = d,
      prior  = base_priors,
      family = gaussian(),
      cores  = 4, iter = 2000, seed = 42,
      file   = sprintf("../results/m_proc_%s", size)
    )
  })
```

```{r proc-brms-coef, fig.height=4}
proc_brms %>%
  imap_dfr(function(m, size) {
    as_draws_df(m) %>%
      select(starts_with("b_"), -b_Intercept) %>%
      pivot_longer(everything(), names_to = "term", values_to = "estimate") %>%
      mutate(model_size = size)
  }) %>%
  mutate(
    model_size = factor(model_size, levels = MODEL_LEVELS),
    term = recode(term,
                  b_log_freq_sc              = "log freq",
                  b_ord_dom_sc               = "ordering dominance",
                  `b_log_freq_sc:ord_dom_sc` = "freq × dom")
  ) %>%
  ggplot(aes(x = estimate, y = term, fill = model_size)) +
  stat_halfeye(position = position_dodge(width = 0.6), alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  scale_fill_brewer(palette = "Set1", name = "Model") +
  labs(x = "Posterior estimate", y = NULL,
       title = "Procrustes distance: Bayesian model coefficients") +
  theme_minimal()
```

```{r proc-cond-effects, fig.height=4}
map(proc_sizes, function(size) {
  ce <- conditional_effects(
    proc_brms[[size]],
    effects = "log_freq_sc:ord_dom_sc"
  )
  plot(ce, plot = FALSE)[[1]] +
    labs(title = size,
         x = "log freq [scaled]", y = "Procrustes dist [scaled]") +
    theme_minimal()
}) %>%
  wrap_plots(nrow = 1) +
  plot_annotation(title = "Conditional effects: frequency × ordering dominance (Procrustes)")
```

---

## In-Context Cosine Similarity (AB vs BA)

Cosine similarity between matched AB and BA phrase representations extracted in
sentence context. High similarity = model treats both orderings alike; low
similarity = ordering substantially changes the representation.

**Predictors:** same as Procrustes — log frequency and ordering dominance.

```{r cos-prep}
cos <- cos_raw %>%
  mutate(model_size = factor(model_size, levels = MODEL_LEVELS)) %>%
  group_by(model, model_size, checkpoint, step, tokens, phrase_AB, layer) %>%
  summarise(cosine_sim = mean(cosine_sim, na.rm = TRUE), .groups = "drop") %>%
  filter_final() %>%
  join_freqs() %>%
  filter(!is.na(overall_freq), overall_freq > 0,
         !is.na(rel_freq), !is.na(cosine_sim)) %>%
  mutate(
    log_freq     = log(overall_freq),
    log_freq_sc  = sc(log_freq),
    ord_dom      = abs(rel_freq - 0.5),
    ord_dom_sc   = sc(ord_dom),
    cos_sc       = sc(cosine_sim),
    log_freq_bin = cut(log_freq, breaks = 3,
                       labels = c("Low freq", "Mid freq", "High freq")),
    ord_dom_bin  = factor(
      case_when(ntile(ord_dom, 3) == 1 ~ "Balanced",
                ntile(ord_dom, 3) == 2 ~ "Moderate",
                ntile(ord_dom, 3) == 3 ~ "Dominant"),
      levels = c("Balanced", "Moderate", "Dominant")
    ),
    model_size   = factor(model_size, levels = MODEL_LEVELS)
  )

cos_sizes <- levels(droplevels(cos$model_size))
cat("Models:", paste(cos_sizes, collapse = ", "), "\n")
cat("N per model:\n"); print(count(cos, model_size))
```

### Visualizations

```{r cos-viz-freq, fig.height=5}
.sm <- fit_smart_smooth(cos, "log_freq", "cosine_sim")
cos %>%
  ggplot(aes(x = log_freq, y = cosine_sim, colour = ord_dom)) +
  geom_point(alpha = 0.4, size = 0.9) +
  geom_ribbon(data = .sm, aes(x = log_freq, ymin = .lwr, ymax = .upr),
              inherit.aes = FALSE, alpha = 0.15, fill = "grey40") +
  geom_line(data = .sm, aes(x = log_freq, y = .fit),
            inherit.aes = FALSE, colour = "black", linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1, scales = "free_x") +
  scale_colour_viridis_c(name = "Ordering dominance\n|RelFreq − 0.5|") +
  labs(x = "log(overall frequency)", y = "Cosine similarity (AB vs BA)",
       title = "In-context cosine similarity vs frequency (final checkpoint, last layer)") +
  theme_minimal() +
  theme(legend.position = "right")
```

```{r cos-viz-ord, fig.height=5}
.sm <- fit_smart_smooth(cos, "ord_dom", "cosine_sim",
                        group_cols = c("model_size", "log_freq_bin"))
cos %>%
  ggplot(aes(x = ord_dom, y = cosine_sim, colour = log_freq_bin)) +
  geom_point(alpha = 0.4, size = 0.9) +
  geom_ribbon(data = .sm, aes(x = ord_dom, ymin = .lwr, ymax = .upr,
                               fill = log_freq_bin),
              inherit.aes = FALSE, alpha = 0.2) +
  geom_line(data = .sm, aes(x = ord_dom, y = .fit, colour = log_freq_bin),
            inherit.aes = FALSE, linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1) +
  scale_colour_brewer(palette = "Blues", aesthetics = c("colour", "fill"),
                      name = "Frequency tertile") +
  labs(x = "Ordering dominance |RelFreq − 0.5|", y = "Cosine similarity (AB vs BA)",
       title = "In-context cosine similarity vs ordering dominance") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r cos-viz-interaction, fig.height=5}
.d  <- cos %>% filter(!is.na(ord_dom_bin))
.sm <- fit_smart_smooth(.d, "log_freq", "cosine_sim",
                        group_cols = c("model_size", "ord_dom_bin"))
.d %>%
  ggplot(aes(x = log_freq, y = cosine_sim,
             colour = ord_dom_bin, fill = ord_dom_bin)) +
  geom_point(alpha = 0.2, size = 0.7) +
  geom_ribbon(data = .sm, aes(x = log_freq, ymin = .lwr, ymax = .upr,
                               fill = ord_dom_bin),
              inherit.aes = FALSE, alpha = 0.2) +
  geom_line(data = .sm, aes(x = log_freq, y = .fit, colour = ord_dom_bin),
            inherit.aes = FALSE, linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1, scales = "free_x") +
  scale_colour_brewer(palette = "Set2", aesthetics = c("colour", "fill"),
                      name = "Ordering dominance") +
  labs(x = "log(overall frequency)", y = "Cosine similarity (AB vs BA)",
       title = "Frequency × ordering dominance interaction (cosine similarity)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### OLS Coefficients

```{r cos-lm-coef, fig.height=4}
cos_sizes %>%
  set_names() %>%
  map_dfr(function(size) {
    lm(cosine_sim ~ log_freq_sc * ord_dom_sc,
       data = cos %>% filter(model_size == size)) %>%
      tidy(conf.int = TRUE) %>%
      filter(term != "(Intercept)")
  }, .id = "model_size") %>%
  mutate(
    model_size = factor(model_size, levels = MODEL_LEVELS),
    term = recode(term,
                  log_freq_sc              = "log freq",
                  ord_dom_sc               = "ordering dominance",
                  `log_freq_sc:ord_dom_sc` = "freq × dom")
  ) %>%
  ggplot(aes(x = estimate, xmin = conf.low, xmax = conf.high,
             y = term, colour = model_size)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_pointrange(position = position_dodge(width = 0.5), linewidth = 0.8) +
  scale_colour_brewer(palette = "Set1", name = "Model") +
  labs(x = "OLS estimate (95% CI)", y = NULL,
       title = "Cosine similarity: OLS coefficients per model") +
  theme_minimal()
```

### brms Models

```{r cos-brms-fit, results="hide"}
cos_brms <- cos_sizes %>%
  set_names() %>%
  map(function(size) {
    d <- cos %>%
      filter(model_size == size) %>%
      mutate(
        cos_sc      = sc(cosine_sim),
        log_freq_sc = sc(log_freq),
        ord_dom_sc  = sc(ord_dom)
      )
    brm(
      cos_sc ~ log_freq_sc * ord_dom_sc,
      data   = d,
      prior  = base_priors,
      family = gaussian(),
      cores  = 4, iter = 2000, seed = 42,
      file   = sprintf("../results/m_cos_%s", size)
    )
  })
```

```{r cos-brms-coef, fig.height=4}
cos_brms %>%
  imap_dfr(function(m, size) {
    as_draws_df(m) %>%
      select(starts_with("b_"), -b_Intercept) %>%
      pivot_longer(everything(), names_to = "term", values_to = "estimate") %>%
      mutate(model_size = size)
  }) %>%
  mutate(
    model_size = factor(model_size, levels = MODEL_LEVELS),
    term = recode(term,
                  b_log_freq_sc              = "log freq",
                  b_ord_dom_sc               = "ordering dominance",
                  `b_log_freq_sc:ord_dom_sc` = "freq × dom")
  ) %>%
  ggplot(aes(x = estimate, y = term, fill = model_size)) +
  stat_halfeye(position = position_dodge(width = 0.6), alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  scale_fill_brewer(palette = "Set1", name = "Model") +
  labs(x = "Posterior estimate", y = NULL,
       title = "In-context cosine similarity: Bayesian model coefficients") +
  theme_minimal()
```

```{r cos-cond-effects, fig.height=4}
map(cos_sizes, function(size) {
  ce <- conditional_effects(
    cos_brms[[size]],
    effects = "log_freq_sc:ord_dom_sc"
  )
  plot(ce, plot = FALSE)[[1]] +
    labs(title = size,
         x = "log freq [scaled]", y = "cosine sim [scaled]") +
    theme_minimal()
}) %>%
  wrap_plots(nrow = 1) +
  plot_annotation(title = "Conditional effects: frequency × ordering dominance (cosine similarity)")
```

---

## Compositional Similarity (LogCosSim)

`LogCosSim = log(cos_AB / cos_BA)` where `cos_AB` is the cosine similarity
between the holistic in-context AB-ordering representation and the mean of
each constituent word embedded independently. Positive LogCosSim = AB ordering
is more compositionally predictable than BA ordering.

**Predictors:**
- `log_freq`: log overall corpus frequency
- `relfreq_c = RelFreq − 0.5`: directional relative frequency (−0.5 = always BA, +0.5 = always AB). Unlike the symmetric metrics above, LogCosSim is directional, so the full signed scale is meaningful.

```{r comp-prep}
comp <- comp_raw %>%
  mutate(model_size = factor(model_size, levels = MODEL_LEVELS)) %>%
  filter_final() %>%
  pivot_wider(names_from = ordering, values_from = mean_cosine_sim,
              names_prefix = "comp_sim_") %>%
  join_freqs() %>%
  filter(!is.na(overall_freq), overall_freq > 0,
         !is.na(rel_freq),
         !is.na(comp_sim_AB), comp_sim_AB > 0,
         !is.na(comp_sim_BA), comp_sim_BA > 0) %>%
  mutate(
    LogCosSim    = log(comp_sim_AB / comp_sim_BA),
    LogCosSim_sc = sc(LogCosSim),
    log_freq     = log(overall_freq),
    log_freq_sc  = sc(log_freq),
    relfreq_c    = rel_freq - 0.5,
    log_freq_bin = cut(log_freq, breaks = 3,
                       labels = c("Low freq", "Mid freq", "High freq")),
    relfreq_bin  = factor(
      case_when(ntile(relfreq_c, 3) == 1 ~ "BA-dominant",
                ntile(relfreq_c, 3) == 2 ~ "Balanced",
                ntile(relfreq_c, 3) == 3 ~ "AB-dominant"),
      levels = c("BA-dominant", "Balanced", "AB-dominant")
    ),
    model_size   = factor(model_size, levels = MODEL_LEVELS)
  )

comp_sizes <- levels(droplevels(comp$model_size))
cat("Models:", paste(comp_sizes, collapse = ", "), "\n")
cat("N per model:\n"); print(count(comp, model_size))
```

### Visualizations

```{r comp-viz-freq, fig.height=5}
.sm <- fit_smart_smooth(comp, "log_freq", "LogCosSim")
comp %>%
  ggplot(aes(x = log_freq, y = LogCosSim, colour = relfreq_c)) +
  geom_point(alpha = 0.4, size = 0.9) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_ribbon(data = .sm, aes(x = log_freq, ymin = .lwr, ymax = .upr),
              inherit.aes = FALSE, alpha = 0.15, fill = "grey40") +
  geom_line(data = .sm, aes(x = log_freq, y = .fit),
            inherit.aes = FALSE, colour = "black", linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1, scales = "free_x") +
  scale_colour_gradient2(
    low = "#2166ac", mid = "grey80", high = "#d6604d", midpoint = 0,
    name = "RelFreq − 0.5\n(−0.5 = BA, +0.5 = AB)"
  ) +
  labs(x = "log(overall frequency)", y = "LogCosSim",
       title = "LogCosSim vs frequency (final checkpoint, last layer)") +
  theme_minimal() +
  theme(legend.position = "right")
```

```{r comp-viz-relfreq, fig.height=5}
.sm <- fit_smart_smooth(comp, "relfreq_c", "LogCosSim",
                        group_cols = c("model_size", "log_freq_bin"))
comp %>%
  ggplot(aes(x = relfreq_c, y = LogCosSim, colour = log_freq_bin)) +
  geom_point(alpha = 0.4, size = 0.9) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_ribbon(data = .sm, aes(x = relfreq_c, ymin = .lwr, ymax = .upr,
                               fill = log_freq_bin),
              inherit.aes = FALSE, alpha = 0.2) +
  geom_line(data = .sm, aes(x = relfreq_c, y = .fit, colour = log_freq_bin),
            inherit.aes = FALSE, linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1) +
  scale_colour_brewer(palette = "Blues", aesthetics = c("colour", "fill"),
                      name = "Frequency tertile") +
  labs(x = "Relative frequency (RelFreq − 0.5)", y = "LogCosSim",
       title = "LogCosSim vs relative frequency") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r comp-viz-interaction, fig.height=5}
.d  <- comp %>% filter(!is.na(relfreq_bin))
.sm <- fit_smart_smooth(.d, "log_freq", "LogCosSim",
                        group_cols = c("model_size", "relfreq_bin"))
.d %>%
  ggplot(aes(x = log_freq, y = LogCosSim,
             colour = relfreq_bin, fill = relfreq_bin)) +
  geom_point(alpha = 0.2, size = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_ribbon(data = .sm, aes(x = log_freq, ymin = .lwr, ymax = .upr,
                               fill = relfreq_bin),
              inherit.aes = FALSE, alpha = 0.2) +
  geom_line(data = .sm, aes(x = log_freq, y = .fit, colour = relfreq_bin),
            inherit.aes = FALSE, linewidth = 0.8) +
  facet_wrap(~model_size, nrow = 1, scales = "free_x") +
  scale_colour_brewer(palette = "RdBu", direction = -1,
                      aesthetics = c("colour", "fill"),
                      name = "Ordering preference") +
  labs(x = "log(overall frequency)", y = "LogCosSim",
       title = "Frequency × relative frequency interaction (LogCosSim)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### OLS Coefficients

```{r comp-lm-coef, fig.height=4}
comp_sizes %>%
  set_names() %>%
  map_dfr(function(size) {
    lm(LogCosSim ~ log_freq_sc * relfreq_c,
       data = comp %>% filter(model_size == size)) %>%
      tidy(conf.int = TRUE) %>%
      filter(term != "(Intercept)")
  }, .id = "model_size") %>%
  mutate(
    model_size = factor(model_size, levels = MODEL_LEVELS),
    term = recode(term,
                  log_freq_sc              = "log freq",
                  relfreq_c                = "relative freq",
                  `log_freq_sc:relfreq_c`  = "freq × relfreq")
  ) %>%
  ggplot(aes(x = estimate, xmin = conf.low, xmax = conf.high,
             y = term, colour = model_size)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_pointrange(position = position_dodge(width = 0.5), linewidth = 0.8) +
  scale_colour_brewer(palette = "Set1", name = "Model") +
  labs(x = "OLS estimate (95% CI)", y = NULL,
       title = "LogCosSim: OLS coefficients per model") +
  theme_minimal()
```

### brms Models

```{r comp-brms-fit, results="hide"}
comp_brms <- comp_sizes %>%
  set_names() %>%
  map(function(size) {
    d <- comp %>%
      filter(model_size == size) %>%
      mutate(
        LogCosSim_sc = sc(LogCosSim),
        log_freq_sc  = sc(log_freq),
        relfreq_c    = rel_freq - 0.5   # fixed theoretical centering, not mean-centered
      )
    brm(
      LogCosSim_sc ~ log_freq_sc * relfreq_c,
      data   = d,
      prior  = base_priors,
      family = gaussian(),
      cores  = 4, iter = 2000, seed = 42,
      file   = sprintf("../results/m_comp2_%s", size)
    )
  })
```

```{r comp-brms-coef, fig.height=4}
comp_brms %>%
  imap_dfr(function(m, size) {
    as_draws_df(m) %>%
      select(starts_with("b_"), -b_Intercept) %>%
      pivot_longer(everything(), names_to = "term", values_to = "estimate") %>%
      mutate(model_size = size)
  }) %>%
  mutate(
    model_size = factor(model_size, levels = MODEL_LEVELS),
    term = recode(term,
                  b_log_freq_sc             = "log freq",
                  b_relfreq_c               = "relative freq",
                  `b_log_freq_sc:relfreq_c` = "freq × relfreq")
  ) %>%
  ggplot(aes(x = estimate, y = term, fill = model_size)) +
  stat_halfeye(position = position_dodge(width = 0.6), alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  scale_fill_brewer(palette = "Set1", name = "Model") +
  labs(x = "Posterior estimate", y = NULL,
       title = "LogCosSim: Bayesian model coefficients") +
  theme_minimal()
```

```{r comp-cond-effects, fig.height=4}
map(comp_sizes, function(size) {
  ce <- conditional_effects(
    comp_brms[[size]],
    effects = "log_freq_sc:relfreq_c"
  )
  plot(ce, plot = FALSE)[[1]] +
    labs(title = size,
         x = "log freq [scaled]", y = "LogCosSim [scaled]") +
    theme_minimal()
}) %>%
  wrap_plots(nrow = 1) +
  plot_annotation(title = "Conditional effects: frequency × relative frequency (LogCosSim)")
```
