---
title: "Binomial Representation Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5)
```

## Setup

```{r packages}
library(tidyverse)
library(brms)
library(tidybayes)
library(patchwork)
```

```{r load-data}
df_raw <- read_csv("../results/binomial_representations.csv")

# Replace web-corpus frequency estimates with counts from the BabyLM training
# corpus (produced by Scripts/get_corpus_freqs.py).
corpus_freqs <- read_csv("../Data/babylm_corpus_freqs.csv") %>%
  select(phrase_AB, corpus_freq = overall_freq, corpus_rel_freq = rel_freq,
         freq_AB, freq_BA)

df_raw <- df_raw %>%
  select(-overall_freq, -rel_freq) %>%
  left_join(corpus_freqs, by = "phrase_AB")

# Each row is one (model × checkpoint × binomial × layer) observation.
# self_sim_AB is always ≤ 0: closer to 0 = more consistent across contexts.
# procrustes_dist is normalised residual of Procrustes(AB → BA): lower = better alignment.
glimpse(df_raw)
```

## Data Preparation

```{r prep}
# Keep only the last layer per model (most processed representation).
# 125m: layer 12, 350m/1.3b: layer 24.
last_layer <- df_raw %>%
  group_by(model_size) %>%
  summarise(last_layer = max(layer), .groups = "drop")

df <- df_raw %>%
  left_join(last_layer, by = "model_size") %>%
  filter(layer == last_layer) %>%
  select(-last_layer) %>%
  mutate(
    log_freq        = log1p(corpus_freq),   # log1p handles any zero-count binomials
    log_tokens      = log(tokens),
    # Zero-corpus binomials get rel_freq = 0.5 (no ordering evidence either way)
    corpus_rel_freq = if_else(is.na(corpus_rel_freq), 0.5, corpus_rel_freq),
    # Negate self_sim so higher = more consistent (easier to interpret)
    consistency = -self_sim_AB,
    model_size  = factor(model_size, levels = c("125m", "350m", "1.3b"))
  )

# Label early / mid / final checkpoints per model
checkpoint_labels <- df %>%
  distinct(model_size, step, tokens) %>%
  group_by(model_size) %>%
  arrange(step) %>%
  mutate(ckpt_label = case_when(
    row_number() == 1              ~ "early",
    row_number() == n()            ~ "final",
    row_number() == round(n() / 2) ~ "mid",
    TRUE                           ~ NA_character_
  )) %>%
  filter(!is.na(ckpt_label)) %>%
  ungroup()

df_labeled <- df %>%
  left_join(checkpoint_labels %>% select(model_size, step, ckpt_label),
            by = c("model_size", "step"))

cat("Rows per checkpoint label:\n")
print(count(df_labeled %>% filter(!is.na(ckpt_label)), model_size, ckpt_label, step))
```

## Visualisations

### Procrustes distance vs consistency at final checkpoint

```{r viz-consistency, fig.height=4}
df_final <- df_labeled %>% filter(ckpt_label == "final")

ggplot(df_final, aes(x = consistency, y = procrustes_dist, colour = model_size)) +
  geom_point(alpha = 0.25, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~model_size, scales = "free_x") +
  scale_colour_brewer(palette = "Set1", guide = "none") +
  labs(
    x     = "Consistency (−self_sim_AB; higher = more consistent)",
    y     = "Procrustes distance",
    title = "Procrustes distance vs representation consistency (final checkpoint, last layer)"
  ) +
  theme_minimal()
```

### Procrustes distance vs overall frequency across checkpoints

```{r viz-freq, fig.height=9}
df_labeled %>%
  filter(!is.na(ckpt_label), corpus_freq > 0) %>%
  mutate(ckpt_label = factor(ckpt_label, levels = c("early", "mid", "final"))) %>%
  ggplot(aes(x = log_freq, y = procrustes_dist, colour = model_size)) +
  geom_point(alpha = 0.25, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(ckpt_label ~ model_size) +
  scale_colour_brewer(palette = "Set1", guide = "none") +
  labs(
    x     = "log(overall frequency)",
    y     = "Procrustes distance",
    title = "Procrustes distance vs corpus frequency (last layer)"
  ) +
  theme_minimal()
```

### Interaction of overall frequency × relative frequency on Procrustes distance

Binomials split into high/low relative frequency (median split within each model),
then Procrustes distance plotted against log overall frequency.
High rel_freq = AB order dominant; low rel_freq = BA order dominant.

```{r viz-freq-relfreq-interaction, fig.height=9}
df_labeled %>%
  filter(!is.na(ckpt_label), corpus_freq > 0, !is.na(corpus_rel_freq)) %>%
  mutate(
    ckpt_label = factor(ckpt_label, levels = c("early", "mid", "final")),
    rel_freq_group = if_else(corpus_rel_freq >= 0.5, "AB-dominant (rel_freq ≥ 0.5)",
                                                      "BA-dominant (rel_freq < 0.5)")
  ) %>%
  ggplot(aes(x = log_freq, y = procrustes_dist,
             colour = rel_freq_group, fill = rel_freq_group)) +
  geom_point(alpha = 0.2, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(ckpt_label ~ model_size) +
  scale_colour_brewer(palette = "Set2", aesthetics = c("colour", "fill")) +
  labs(
    x      = "log(overall frequency)",
    y      = "Procrustes distance",
    colour = "Ordering preference",
    fill   = "Ordering preference",
    title  = "Procrustes distance: overall freq × relative freq interaction (last layer)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### Predictor correlation: consistency vs frequency

```{r viz-pred-corr, fig.height=4}
ggplot(df_final, aes(x = log_freq, y = consistency, colour = model_size)) +
  geom_point(alpha = 0.2, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~model_size, scales = "free_y") +
  scale_colour_brewer(palette = "Set1", guide = "none") +
  labs(
    x = "log(overall frequency)",
    y = "Consistency (−self_sim_AB)",
    title = "Predictor correlation: consistency vs frequency"
  ) +
  theme_minimal()
```

### Effect of frequency on Procrustes distance across layers

Spearman correlation between log corpus frequency and Procrustes distance
computed at each layer, at the final checkpoint. Positive = higher-frequency
binomials have higher Procrustes distance (more dissimilar ordering representations);
negative = higher-frequency binomials have more similar ordering representations.

```{r viz-freq-by-layer, fig.height=4}
# Use full dataset (all layers), filtered to final checkpoint per model
final_steps <- checkpoint_labels %>% filter(ckpt_label == "final") %>%
  select(model_size, step)

df_all_layers_final <- df_raw %>%
  semi_join(final_steps, by = c("model_size", "step")) %>%
  mutate(
    log_freq   = log1p(corpus_freq),
    model_size = factor(model_size, levels = c("125m", "350m", "1.3b")),
    layer_prop = layer / max(layer),   # normalise within each row (max varies by model)
  )

# Recompute per-model layer_prop correctly
df_all_layers_final <- df_all_layers_final %>%
  group_by(model_size) %>%
  mutate(layer_prop = layer / max(layer)) %>%
  ungroup()

# Spearman correlation between log_freq and procrustes_dist at each layer × model
freq_layer_cor <- df_all_layers_final %>%
  group_by(model_size, layer, layer_prop) %>%
  summarise(
    r = cor(log_freq, procrustes_dist, method = "spearman", use = "complete.obs"),
    .groups = "drop"
  )

ggplot(freq_layer_cor, aes(x = layer_prop, y = r, colour = model_size)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  scale_colour_brewer(palette = "Set1") +
  scale_x_continuous(labels = scales::percent) +
  labs(
    x      = "Layer (% of total depth)",
    y      = "Spearman r (log freq vs Procrustes distance)",
    title  = "Effect of corpus frequency on Procrustes distance across layers (final checkpoint)",
    colour = "Model"
  ) +
  theme_minimal()
```

### Procrustes distance across training

```{r viz-training, fig.height=5}
df_avg <- df %>%
  group_by(model_size, log_tokens) %>%
  summarise(mean_proc = mean(procrustes_dist),
            se_proc   = sd(procrustes_dist) / sqrt(n()),
            .groups   = "drop")

ggplot(df_avg, aes(x = log_tokens, y = mean_proc,
                   colour = model_size, fill = model_size)) +
  geom_ribbon(aes(ymin = mean_proc - se_proc,
                  ymax = mean_proc + se_proc), alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.8) +
  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x      = "log(tokens seen)",
    y      = "Mean Procrustes distance",
    title  = "Procrustes distance over training (last layer, mean ± SE across binomials)",
    colour = "Model", fill = "Model"
  ) +
  theme_minimal()
```

## brms Models

All models are fit separately per LLM (125m, 350m, 1.3b).

Outcome: `procrustes_dist`.
Predictors: `consistency` (−self_sim_AB) and `log_freq`, both scaled within each model fit.
Random effect: `(1 | phrase_AB)`.
Family: Gaussian (beta regression would be more principled for a 0–1 outcome but is a reasonable starting point).

```{r brms-priors}
priors_base <- c(
  prior(normal(0, 1),   class = b),
  prior(normal(0.5, 1), class = Intercept),
  prior(exponential(1), class = sd),
  prior(exponential(1), class = sigma)
)

model_sizes <- c("125m", "350m", "1.3b")
```

### Checkpoint-specific models (early / mid / final)

One model per LLM per timepoint — 9 models total, cached to disk.

```{r brms-checkpoint-models, results="hide"}
formula_base <- procrustes_dist ~ cons_sc * freq_sc + (1 | phrase_AB)

# Returns a named list: list(early = m, mid = m, final = m) for one model size
fit_ckpt_models <- function(size) {
  map(c("early", "mid", "final"), function(label) {
    d <- df_labeled %>%
      filter(model_size == size, ckpt_label == label) %>%
      mutate(cons_sc = as.vector(scale(consistency)),
             freq_sc = as.vector(scale(log_freq)))

    brm(
      formula_base,
      data   = d,
      prior  = priors_base,
      family = gaussian(),
      cores  = 4,
      iter   = 2000,
      seed   = 42,
      file   = sprintf("../results/m_%s_%s", size, label)
    )
  }) %>% set_names(c("early", "mid", "final"))
}

ckpt_models <- map(model_sizes, fit_ckpt_models) %>% set_names(model_sizes)
```

```{r checkpoint-summaries}
iwalk(ckpt_models, function(size_models, size) {
  iwalk(size_models, function(m, label) {
    cat(sprintf("\n=== %s — %s checkpoint ===\n", size, label))
    print(summary(m))
  })
})
```

```{r checkpoint-coef-plot, fig.height=8}
coef_df <- imap_dfr(ckpt_models, function(size_models, size) {
  imap_dfr(size_models, function(m, label) {
    as_draws_df(m) %>%
      select(starts_with("b_"), -b_Intercept) %>%
      pivot_longer(everything(), names_to = "term", values_to = "estimate") %>%
      mutate(checkpoint = label, model_size = size)
  })
}) %>%
  mutate(
    checkpoint = factor(checkpoint, levels = c("early", "mid", "final")),
    model_size = factor(model_size, levels = model_sizes)
  )

ggplot(coef_df, aes(x = estimate, y = term, fill = checkpoint)) +
  stat_halfeye(position = position_dodge(width = 0.6), alpha = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  facet_wrap(~model_size, ncol = 1) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    x     = "Posterior estimate",
    y     = NULL,
    title = "Coefficients across early / mid / final checkpoints, per LLM",
    fill  = "Checkpoint"
  ) +
  theme_minimal()
```

### Full model: effect of training progress (all checkpoints)

```{r brms-full-models, results="hide"}
formula_full <- procrustes_dist ~ cons_sc * freq_sc * log_tokens_sc + (1 | phrase_AB)

fit_full_model <- function(size) {
  d <- df %>%
    filter(model_size == size) %>%
    mutate(cons_sc       = as.vector(scale(consistency)),
           freq_sc       = as.vector(scale(log_freq)),
           log_tokens_sc = as.vector(scale(log_tokens)))

  brm(
    formula_full,
    data   = d,
    prior  = c(
      prior(normal(0, 1),   class = b),
      prior(normal(0.5, 1), class = Intercept),
      prior(exponential(1), class = sd),
      prior(exponential(1), class = sigma)
    ),
    family = gaussian(),
    cores  = 4,
    iter   = 2000,
    seed   = 42,
    file   = sprintf("../results/m_%s_full", size)
  )
}

full_models <- map(model_sizes, fit_full_model) %>% set_names(model_sizes)
```

```{r full-summaries}
iwalk(full_models, function(m, size) {
  cat(sprintf("\n=== %s — full model ===\n", size))
  print(summary(m))
})
```

```{r full-coef-plot, fig.height=7}
imap_dfr(full_models, function(m, size) {
  as_draws_df(m) %>%
    select(starts_with("b_"), -b_Intercept) %>%
    pivot_longer(everything(), names_to = "term", values_to = "estimate") %>%
    mutate(model_size = size)
}) %>%
  mutate(model_size = factor(model_size, levels = model_sizes)) %>%
  ggplot(aes(x = estimate, y = term)) +
  stat_halfeye(fill = "#4393c3", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  facet_wrap(~model_size, ncol = 1) +
  labs(
    x     = "Posterior estimate",
    y     = NULL,
    title = "Full model coefficients (all checkpoints), per LLM"
  ) +
  theme_minimal()
```

### Conditional effects: consistency × frequency (final checkpoint)

```{r cond-effects, fig.height=4}
plots <- map(model_sizes, function(size) {
  ce <- conditional_effects(
    ckpt_models[[size]][["final"]],
    effects    = "cons_sc:freq_sc",
    re_formula = NA
  )
  plot(ce, plot = FALSE)[[1]] +
    labs(title = size, x = "Consistency (scaled)", y = "Procrustes distance") +
    theme_minimal()
})

wrap_plots(plots, nrow = 1) +
  plot_annotation(title = "Consistency × frequency interaction at final checkpoint")
```
