---
title: "Binomial Representation Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5)
```

## Setup

```{r packages}
library(tidyverse)
library(brms)
library(tidybayes)
library(patchwork)
```

```{r load-data}
df_raw <- read_csv("../results/binomial_representations.csv")

# Each row is one (model × checkpoint × binomial × layer) observation.
# self_sim_AB is always ≤ 0: closer to 0 = more consistent across contexts.
# procrustes_dist is normalised residual of Procrustes(AB → BA): lower = better alignment.
glimpse(df_raw)
```

## Data Preparation

```{r prep}
# Keep only the last layer per model (most processed representation).
# 125m: layer 12, 350m/1.3b: layer 24.
last_layer <- df_raw %>%
  group_by(model_size) %>%
  summarise(last_layer = max(layer), .groups = "drop")

df <- df_raw %>%
  left_join(last_layer, by = "model_size") %>%
  filter(layer == last_layer) %>%
  select(-last_layer) %>%
  mutate(
    log_freq    = log(overall_freq),
    log_tokens  = log(tokens),
    # Negate self_sim so higher = more consistent (easier to interpret)
    consistency = -self_sim_AB,
    model_size  = factor(model_size, levels = c("125m", "350m", "1.3b"))
  )

# Label early / mid / final checkpoints per model
checkpoint_labels <- df %>%
  distinct(model_size, step, tokens) %>%
  group_by(model_size) %>%
  arrange(step) %>%
  mutate(ckpt_label = case_when(
    row_number() == 1          ~ "early",
    row_number() == n()        ~ "final",
    row_number() == round(n() / 2) ~ "mid",
    TRUE                       ~ NA_character_
  )) %>%
  filter(!is.na(ckpt_label)) %>%
  ungroup()

df_labeled <- df %>%
  left_join(checkpoint_labels %>% select(model_size, step, ckpt_label),
            by = c("model_size", "step"))

cat("Rows per checkpoint label:\n")
print(count(df_labeled %>% filter(!is.na(ckpt_label)), model_size, ckpt_label, step))
```

## Visualisations

### Procrustes distance vs consistency at final checkpoint

```{r viz-consistency, fig.height=4}
df_final <- df_labeled %>% filter(ckpt_label == "final")

p1 <- ggplot(df_final, aes(x = consistency, y = procrustes_dist, colour = model_size)) +
  geom_point(alpha = 0.25, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    x     = "Consistency (−self_sim_AB; higher = more consistent)",
    y     = "Procrustes distance",
    title = "Procrustes distance vs representation consistency (final checkpoint, last layer)",
    colour = "Model"
  ) +
  theme_minimal()
p1
```

### Procrustes distance vs overall frequency at final checkpoint

```{r viz-freq, fig.height=4}
p2 <- ggplot(df_final, aes(x = log_freq, y = procrustes_dist, colour = model_size)) +
  geom_point(alpha = 0.25, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    x     = "log(overall frequency)",
    y     = "Procrustes distance",
    title = "Procrustes distance vs corpus frequency (final checkpoint, last layer)",
    colour = "Model"
  ) +
  theme_minimal()
p2
```

### Consistency vs frequency (to understand predictor correlation)

```{r viz-pred-corr, fig.height=4}
p3 <- ggplot(df_final, aes(x = log_freq, y = consistency, colour = model_size)) +
  geom_point(alpha = 0.2, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    x = "log(overall frequency)",
    y = "Consistency (−self_sim_AB)",
    title = "Predictor correlation: consistency vs frequency"
  ) +
  theme_minimal()
p3
```

### Procrustes distance across training

```{r viz-training, fig.height=5}
df_avg <- df %>%
  group_by(model_size, log_tokens, tokens) %>%
  summarise(mean_proc = mean(procrustes_dist),
            se_proc   = sd(procrustes_dist) / sqrt(n()),
            .groups = "drop")

p4 <- ggplot(df_avg, aes(x = log_tokens, y = mean_proc,
                          colour = model_size, fill = model_size)) +
  geom_ribbon(aes(ymin = mean_proc - se_proc,
                  ymax = mean_proc + se_proc), alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.8) +
  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x      = "log(tokens seen)",
    y      = "Mean Procrustes distance",
    title  = "Procrustes distance over training (last layer, mean ± SE across binomials)",
    colour = "Model", fill = "Model"
  ) +
  theme_minimal()
p4
```

## brms Models

Outcome: `procrustes_dist` (normalised Procrustes residual, 0–1).
Predictors: `consistency` (−self_sim_AB) and `log_freq`, both scaled.
Random effect: `(1 | phrase_AB)` for between-binomial variation.

All predictors are scaled within each model fit so coefficients are in
comparable units. Model family: Gaussian (a beta regression would be more
appropriate for a bounded outcome but Gaussian is a reasonable starting point).

```{r brms-priors}
priors_base <- c(
  prior(normal(0, 1),   class = b),
  prior(normal(0.5, 1), class = Intercept),
  prior(exponential(1), class = sd),
  prior(exponential(1), class = sigma)
)
```

### Models at early, mid, and final checkpoints

```{r brms-checkpoint-models, results="hide"}
make_ckpt_df <- function(label) {
  df_labeled %>%
    filter(ckpt_label == label) %>%
    mutate(
      cons_sc  = as.vector(scale(consistency)),
      freq_sc  = as.vector(scale(log_freq))
    )
}

df_early <- make_ckpt_df("early")
df_mid   <- make_ckpt_df("mid")
df_final_sc <- make_ckpt_df("final")

formula_base <- procrustes_dist ~ cons_sc * freq_sc + model_size + (1 | phrase_AB)

m_early <- brm(
  formula_base,
  data    = df_early,
  prior   = priors_base,
  family  = gaussian(),
  cores   = 4,
  iter    = 2000,
  seed    = 42,
  file    = "../results/m_early"
)

m_mid <- brm(
  formula_base,
  data    = df_mid,
  prior   = priors_base,
  family  = gaussian(),
  cores   = 4,
  iter    = 2000,
  seed    = 42,
  file    = "../results/m_mid"
)

m_final <- brm(
  formula_base,
  data    = df_final_sc,
  prior   = priors_base,
  family  = gaussian(),
  cores   = 4,
  iter    = 2000,
  seed    = 42,
  file    = "../results/m_final"
)
```

```{r checkpoint-summaries}
cat("=== Early checkpoint ===\n");  print(summary(m_early))
cat("\n=== Mid checkpoint ===\n");  print(summary(m_mid))
cat("\n=== Final checkpoint ===\n"); print(summary(m_final))
```

```{r checkpoint-coef-plot, fig.height=6}
tidy_model <- function(m, label) {
  as_draws_df(m) %>%
    select(starts_with("b_")) %>%
    pivot_longer(everything(), names_to = "term", values_to = "estimate") %>%
    filter(term != "b_Intercept") %>%
    mutate(checkpoint = label)
}

coef_df <- bind_rows(
  tidy_model(m_early, "early"),
  tidy_model(m_mid,   "mid"),
  tidy_model(m_final, "final")
) %>%
  mutate(checkpoint = factor(checkpoint, levels = c("early", "mid", "final")))

ggplot(coef_df, aes(x = estimate, y = term, fill = checkpoint)) +
  stat_halfeye(position = position_dodge(width = 0.6), alpha = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    x     = "Posterior estimate",
    y     = "Term",
    title = "Coefficients across early / mid / final checkpoints",
    fill  = "Checkpoint"
  ) +
  theme_minimal()
```

### Full model: effect of training progress

Uses all checkpoints. `log_tokens_sc` captures training progress on a log scale
(comparable across model sizes).

```{r brms-full-model, results="hide"}
df_full <- df %>%
  mutate(
    cons_sc       = as.vector(scale(consistency)),
    freq_sc       = as.vector(scale(log_freq)),
    log_tokens_sc = as.vector(scale(log_tokens))
  )

m_full <- brm(
  procrustes_dist ~ cons_sc * freq_sc * log_tokens_sc + model_size + (1 | phrase_AB),
  data   = df_full,
  prior  = c(
    prior(normal(0, 1),   class = b),
    prior(normal(0.5, 1), class = Intercept),
    prior(exponential(1), class = sd),
    prior(exponential(1), class = sigma)
  ),
  family = gaussian(),
  cores  = 4,
  iter   = 2000,
  seed   = 42,
  file   = "../results/m_full"
)
```

```{r full-summary}
summary(m_full)
```

```{r full-coef-plot, fig.height=5}
as_draws_df(m_full) %>%
  select(starts_with("b_")) %>%
  pivot_longer(everything(), names_to = "term", values_to = "estimate") %>%
  filter(term != "b_Intercept") %>%
  ggplot(aes(x = estimate, y = term)) +
  stat_halfeye(fill = "#4393c3", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  labs(
    x     = "Posterior estimate",
    y     = "Term",
    title = "Full model coefficients (all checkpoints)"
  ) +
  theme_minimal()
```

### Conditional effects: consistency × frequency interaction

```{r cond-effects-final, fig.height=4}
ce <- conditional_effects(
  m_final,
  effects  = "cons_sc:freq_sc",
  re_formula = NA
)
plot(ce, plot = FALSE)[[1]] +
  labs(
    x     = "Consistency (scaled)",
    y     = "Procrustes distance",
    title = "Consistency × frequency interaction (final checkpoint)",
    colour = "log(freq) (scaled)",
    fill   = "log(freq) (scaled)"
  ) +
  theme_minimal()
```
